Demography and selection since maize domestication

### 2015-06-04 manuscript.tex
update

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
%\usepackage{hyperref}
\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }

@@

\dropcap{D}omesticated plant species evolve in a unique fashion
compared to their wild relatives \textcolor{red}{(ref here)}. This
is a result of both the anthropomorphic nature of artificial selection on
domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
bottleneck that they tend to have experienced
\cite{ross2007}. However, the
complex interplay between selective pressures and demographic limitations in
domesticated species is not well understood. Despite the fact that,
researchers have searched the genomes of several species for evidence
of positive selection during domestication
\textcolor{red}{(citations here)}, these types of studies tend to focus on
identifying or mapping particular genes or regions that play an
important role in phenotypic evolution. Our knowledge regarding the impacts of demography and
selection have on whole-genome patterns of genetic variability therefore remain limited.

\textcolor{red}{
This is the beginning of the article. For now, I will give a
bulleted list of the introduction:
\begin{enumerate}
\item Plant domestication
\item Maize
\item Maize domestication
\item Importance of demography
\item Background selection
\end{enumerate}
}

>>>>

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
%\usepackage{hyperref}
\usepackage[normalem]{ulem}
\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }

@@

\dropcap{D}omesticated plant species evolve in a unique fashion
compared to their wild relatives \textcolor{red}{(ref here)}. This
is a result of both the anthropomorphic nature of artificial selection on
domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
bottleneck(s) that they tend to have experienced
\cite{ross2007}. However, the
complex interplay between selective pressures and demographic
limitations, and the impact that this interplay has on identifying
selection and understanding demography, is not fully understood. Although a large body of
research that involves searching the genomes of domesticated species for evidence
of positive selection exists \cite{hufford2012, he2011, OTHERS?}, these studies tend to focus on
identifying or mapping particular genes or regions that play an
important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
selection have on whole-genome patterns of genetic variability therefore remains limited.

For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
parameters of a population such as effective population size, structure, and expansion
history \cite{luikart2003, gutenkunst2009}. However, researchers have
called into question whether or not there are reasonable approaches
for identifying neutral regions of the genome, since the effects of
selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
\emph{Drosophila} it appears that the majority of the genome is
impacted by the effects of selection through linkage
\cite{sella2009}. This observation begs the question, how can
demographic parameters be estimated independently of
selective parameters? Human researchers have attempted to address this
problem by limiting analyses to only sites far from genes
\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
difficult to be certain that sites even in gene-poor regions of the
genome are not influenced by linked selection. Ultimately, an understanding
of which sites have the potential to be adaptive, and how these affect
genome-wide patterns of variability through linkage, is required for
reasonable demographic inference.

Maize represents an excellent organism to study these
phenomena. Maize is a species of tremendous importance worldwide as
both a staple crop \cite{SOMETHING} and as a model for
understanding crop evolution \cite{SOMETHINGS}. Broadly speaking, archaeological and genetic studies have
established that maize domestication is likely to have taken place in
Central Mexico approximately 9,000 years bp
\cite{SOMETHING,SOMETHING}. Teosinte, the most
recent wild ancestor to maize, remains extant throughout much of the
Americas \cite{SOMETHING}. Additionally, several large-effect
domestication loci \cite{SOMETHINGS} and putative domestication
regions \cite{hufford2012} have been identified. But despite all that is
known about maize domestication, the parameters of the
domestication process remain uncertain. Specifically, the size of the
maize domestication bottleneck has not been estimated independently of
the bottleneck's duration, nor are there sequence-based estimates of the effective
population size of modern maize. Sequence information from maize and
teosinte plants may therefore be utilized to address these questions.

To that end, the objectives of our study were to 1) investigate the
relative importance of different forms of selection on whole-genome
variability in both maize and teosinte, as different forms of selection affect DNA
variability remarkably differently; 2) research the impact that the
domestication process has had on genetic variability in maize, and how
this compares to the impact of a different demographic history in
teosinte; and 3) precisely estimate the parameters of the
maize domestication bottleneck. We show that our third objective, estimating the
parameters of domestication, is not possible without first completing
 objectives one and two, which demonstrate that as in humans \cite{gazave2014}, the
majority of maize non-genic DNA may reasonably be treated as
neutral. We achieve these objectives by utilizing
whole-genome-sequence information from 23 maize and 13 teosinte lines
sequenced as part of the Maize HapMap 2 panel \cite{chia2012}.
